Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-12-02 analysis/paper/paper.qmd
Edited plots based on the look in Wordfile.

ggplot(tcsa,
        aes(0.3, 
            min_tcsa)) +
  # show regions for tools from Lomabrd 2021 & 2022
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
  # https://assets.researchsquare.com/files/rs-2016964/v1/00d381a3-3ecf-4d5b-b2ec-89a41227bbd5.pdf?c=1662475267
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 ) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42 ) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing spear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 1.1) +
  coord_cartesian(xlim = c(-0.1,1)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")
```

@@

```{r}
#| label: fig-tcsa-raw-materials
#| fig-cap: "TCSA values by lithic raw material."
#| fig-width:   4
#| fig-height:  3.5

@@

 ggplot(tcsa_raw,
              aes(y = min_tcsa, 
                  x = label_raw)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Raw materials", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 6)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 1.9) +
   annotate("rect", 
           xmin = 6.4,
           xmax = 6.5,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 6.4,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 6.5,
           xmax = 6.6,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 6.5,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 6.6,
           xmax = 6.7,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = 6.6,
           y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 6.7,
           xmax = 6.8,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = 6.7,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 6.8,
           xmax = 6.9,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing\nspear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = 6.8,
           y = (200-80) /2 + 80 )

@@

plot_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(1, 
            `Weight(g)`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.4,
                   size = 1.5) +
  labs(shape='Cluster',
       y = "Weight (g)") +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        )

plot_tcsa_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(0.3, min_tcsa)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.4,
                   size = 1.5) +
  labs(shape='Cluster') +
  theme_bw(base_size = 6) +
  # show regions for tools from Table 3 of Lomabrd 2021
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing\nspear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  coord_cartesian(xlim = c(-0.1,1.2)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")

@@

```{r}
#| label: fig-tcsa-radio-carbon
#| fig-cap: "A: Distribution of TCSA values over time. The light gray shade indicates Marine isotope stages (MIS) 3, the gray shade indicates MIS 2 and the darkest shade indicates the duration of LGM. B: Distribution of TCSA values grouped by LGM event"
#| fig-width:   4.5
#| fig-height:  4

@@

# box plot for each MIS stage
three_time <-  ggplot(tcsa_site_ass_mis, 
       aes(y = min_tcsa,
           x = factor(mis_stage, level = c('After', 'LGM', 'Before')))) +
   geom_boxplot() +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "Time Period") +
  geom_quasirandom(alpha = 0.4,
                   size = 2.3) +
    scale_x_discrete(limits=rev) +
     annotate("rect", 
           xmin = 3.4,
           xmax = 3.5,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 3.4,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 3.5,
           xmax = 3.6,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned\narrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = 3.5,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 3.6,
           xmax = 3.7,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = 3.6,
           y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 3.7,
           xmax = 3.8,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = 3.7,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 3.8,
           xmax = 3.9,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing\nspear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = 3.8,
           y = (200-80) /2 + 80 )

plot_grid(all_time, 
          three_time, 
          labels = c('A', 'B'),
          ncol = 1)

>>>>

ggplot(tcsa,
        aes(0.3, 
            min_tcsa)) +
  # show regions for tools from Lomabrd 2021 & 2022
  # https://www-sciencedirect-com.offcampus.lib.washington.edu/science/article/pii/S0277379121002286
  # https://assets.researchsquare.com/files/rs-2016964/v1/00d381a3-3ecf-4d5b-b2ec-89a41227bbd5.pdf?c=1662475267
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 - 5) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42 ) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing spear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  geom_boxplot(alpha = 0.1) +
  geom_quasirandom(alpha = 0.4,
                   size = 0.7) +
  coord_cartesian(xlim = c(-0.1,1)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")
```

@@

```{r}
#| label: fig-tcsa-raw-materials
#| fig-cap: "TCSA values by lithic raw material."
#| fig-width:   4.4
#| fig-height:  3.2

@@

 ggplot(tcsa_raw,
              aes(y = min_tcsa, 
                  x = label_raw)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Raw materials", 
       y = "TCSA") +
  theme(axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        text = element_text(size=8)) +
  scale_x_discrete(guide = guide_axis(angle = 45))+
  geom_quasirandom(alpha = 0.4,
                   size = 1.3) +
   annotate("rect", 
           xmin = 0.5,
           xmax = 1.7,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 1.7,
           xmax = 2.9,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 2.9,
           xmax = 4.1,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("rect", 
           xmin = 4.1,
           xmax = 5.3,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
     annotate("rect", 
           xmin = 5.3,
           xmax = 6.5,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) 
 # + annotate("text", 
 #               size = 2,
 #               label = "Poisoned arrow tips",
 #               hjust = 0,
 #               lineheight = 0.75,
 #           x = 0.75,
   #         y = (18-4) /2 + 4) +
   # annotate("text",
   #          size = 2,
   #             label = "Un-poisoned arrow tips",
   #             hjust = 0,
   #             lineheight = 0.75,
   #         x = 1.85,
   #         y = (47-17) /2 + 17) +
   # annotate("text", 
   #             size = 2,
   #             label = "Dart tips",
   #             hjust = 0,
   #         x = 3.3,
   #         y = (76-40) /2 + 40) +
   # annotate("text", 
   #             size = 2,
   #             label = "Javelin tips",
   #             hjust = 0,
   #         x = 4.4,
   #         y = (90-42) /2 + 42) +
    # annotate("text", 
    #            label = "Stabbing spear tips",
    #            size = 2,
    #            hjust = 0,
    #            lineheight = 0.75,
    #        x = 5.52,
    #        y = (200-80) /2 + 80 )

@@

plot_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(1, 
            `Weight(g)`)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.4,
                   size = 1.1) +
  labs(shape='Cluster',
       y = "Weight (g)") +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        )

plot_tcsa_mass_cluster <- 
ggplot (tcsa_mass_no_na,
        aes(0.3, min_tcsa)) +
  geom_boxplot(outlier.shape = NA) +
  geom_quasirandom(aes(shape = mass_cluster),
                   alpha = 0.4,
                   size = 1.1) +
  labs(shape='Cluster') +
  theme_bw(base_size = 6) +
    annotate("rect", 
           xmin = 0.1,
           xmax = 0.2,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 0.2,
           xmax = 0.3,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Un-poisoned arrow tips",
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 0.3,
           xmax = 0.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Dart tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (76-40) /2 + 40 - 5) +
      annotate("rect", 
           xmin = 0.4,
           xmax = 0.5,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      annotate("text", 
               size = point_function_group_label_size,
               label = "Javelin tips",
               hjust = 0,
           x = point_function_group_label_y,
           y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 0.5,
           xmax = 0.6,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) +
      annotate("text", 
               label = "Stabbing spear tips",
               size = point_function_group_label_size,
               hjust = 0,
               lineheight = 0.75,
           x = point_function_group_label_y,
           y = (200-80) /2 + 80 ) +
  coord_cartesian(xlim = c(-0.1,1.2)) +
  theme_bw(base_size = 6) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
        ) +
  labs(y = "TCSA")

@@

```{r}
#| label: fig-tcsa-radio-carbon
#| fig-cap: "A: Distribution of TCSA values over time. The light gray shade indicates Marine isotope stages (MIS) 3, the gray shade indicates MIS 2 and the darkest shade indicates the duration of LGM. B: Distribution of TCSA values grouped by LGM event"
#| fig-width:   3.8
#| fig-height:  4

@@

# box plot for each MIS stage
three_time <-  ggplot(tcsa_site_ass_mis, 
       aes(y = min_tcsa,
           x = factor(mis_stage, level = c('After', 'LGM', 'Before')))) +
   geom_boxplot() +
   theme_minimal() +
   theme(legend.position="none") +
   labs(y = "TCSA", 
        x = "Time Period") +
  geom_quasirandom(alpha = 0.4,
                   size = 1.9) +
    scale_x_discrete(limits=rev) +
     annotate("rect", 
           xmin = 0.3,
           xmax = 1,
           ymin = 4,
           ymax = 18,
           fill = "green",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.4,
      #      y = (18-4) /2 + 4) +
      annotate("rect", 
           xmin = 1,
           xmax = 1.7,
           ymin = 17,
           ymax = 47,
           fill = "blue",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Un-poisoned\narrow tips",
      #          hjust = 0,
      #          lineheight = 0.75,
      #      x = 3.5,
      #      y = (47-17) /2 + 17) +
      annotate("rect", 
           xmin = 1.7,
           xmax = 2.4,
           ymin = 40,
           ymax = 76,
           fill = "orange",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Dart tips",
      #          hjust = 0,
      #      x = 3.6,
      #      y = (76-40) /2 + 40) +
      annotate("rect", 
           xmin = 2.4,
           xmax = 3.1,
           ymin = 42,
           ymax = 90,
           fill = "purple",
           alpha = 0.2) +
      # annotate("text", 
      #          size = point_function_group_label_size,
      #          label = "Javelin tips",
      #          hjust = 0,
      #      x = 3.7,
      #      y = (90-42) /2 + 42) +
     annotate("rect", 
           xmin = 3.1,
           xmax = 3.8,
           ymin = 80,
           ymax = 200,
           fill = "yellow",
           alpha = 0.2) 
# + annotate("text", 
#                label = "Stabbing\nspear tips",
#                size = point_function_group_label_size,
           #     hjust = 0,
           #     lineheight = 0.75,
           # x = 3.8,
           # y = (200-80) /2 + 80 )

plot_grid(all_time, 
          three_time, 
          labels = c('A', 'B'),
          ncol = 1,
          rel_heights = c(2.7, 1.5))
